The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach

### 2023-07-12 analysis/paper/paper.qmd
Update text and make small corrections in code chunks

In order to describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved over the years of the program. To that end, we employed a text mining approach in which we analysed its output: `r nrow(all_theses)` theses with an average of `r mean(str_length(emu$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(all_theses$grad_year, na.rm = TRUE), "-", max(all_theses$grad_year, na.rm = TRUE))`.

@@

NLP... and the focus on its subtopic topic modeling

@@

Of the total of `r nrow(all_theses)` theses published in the EMU program, full-text documents were available and permissions were granted for `r nrow(emu)` <!--# update after removing theses without permission -->. Metadata was downloaded from the TU Delft Education Repository or obtained from the EMU program coordination records when unavailable on the repository. The full-text documents were obtained either from the TU Delft Education Repository (an API for downloading a large set of full-text data is not available), when available, or from the authors via email. Even though repository entries are open access, it was unclear if the content of the theses can be mined. Considering the relatively small size of the corpus and the possibility of directly contacting the alumni network, permissions were requested individually. Only PDFs for which permission was obtained were included in the corpus.With the exception of a bilingual English-Spanish thesis, all thesis were written in English. The theses are available in PDF format with complex layouts typical to an urbanism project in which text and various types of graphics are combined. @tbl-data gives an overview of the data included in the analysis.

@@

@@

```{r topic-model-lda-1}
#| dependson: short-stop-words

@@

```{r topic-model-lda-2}

# What is the dominant topic in each document?
N = 20
topic_prop_examples <- theta[1:N,]
colnames(topic_prop_examples) <- topic_names
# TODO change to pivot_longer
vizDataFrame <- melt(cbind(data.frame(topic_prop_examples), 
                           document = factor(rownames(topic_prop_examples))), 
                     variable.name = "topic", id.vars = "document")  
ggplot(data = vizDataFrame, aes(topic, value, fill = document), ylab = "proportion") + 
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +  
  coord_flip() +
  facet_wrap(~ document, ncol = 5) +
  theme(legend.position = "none")

@@

# Extract topics as mixtures of terms from LDA model
beta_tidy <- tidy(emu_lda, matrix = "beta")

# The top n most common words in each topic
emu_theses_top_terms <- get_top_words_per_topic(beta_tidy, 20)

@@

```{r}
## Trigrams
emu_trigrams <- emu |> 
  get_ngrams(n = 3, title_col = "title", text_col = "text")

@@

freq_by_rank |> 
  ggplot(aes(rank, `term frequency`, color = title)) +
  geom_line(size = 1.1, alpha = 0.8, show.legend = FALSE) +
  geom_abline(intercept = -1.44, slope = -0.7, color = "gray50", linetype = 2) +
  scale_x_log10() +
  scale_y_log10()

## Visualise the top n words with the highest tf-idf values for each of a subset of theses
emu_words_count |> 
  filter(title == unique(title)[1:6]) |> 
  arrange(desc(tf_idf)) |> 
  mutate(word = factor(word, levels = rev(unique(word)))) |> 
  group_by(title) |> 
  top_n(15) |> 
  ungroup() |> 
  ggplot(aes(word, tf_idf, fill = title)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  facet_wrap(~ title,  ncol = 3, scales = "free") +
  coord_flip()
```

@@

# Summarise word embeddings into document embeddings for modeling purposes
word_matrix <- tidy_emu |> 
  count(title, word) |> 
  cast_sparse(title, word, n)

>>>>

In order to describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved over the years of the program. To that end, we employed a text mining approach in which we analysed its output: 96 theses with an average of `r mean(str_length(emu$text_clean), na.rm = TRUE)` words produced over the years for the duration of the program between `r paste0(min(emu$graduation_year, na.rm = TRUE), "-", max(emu$graduation_year, na.rm = TRUE))`.

@@

In Natural Language Processing (NLP) ... and the focus on its subtopic topic modeling

@@

Of the total of 96 theses published in the EMU program, full-text documents were available and permissions were granted for `r nrow(emu)` <!--# update after removing theses without permission -->. Metadata was downloaded from the TU Delft Education Repository or obtained from the EMU program coordination records when unavailable on the repository. The full-text documents were obtained either from the TU Delft Education Repository (an API for downloading a large set of full-text data is not available), when available, or from the authors via email. Even though repository entries are open access, it was unclear if the content of the theses can be mined. Considering the relatively small size of the corpus and the possibility of directly contacting the alumni network, permissions were requested individually. Only PDFs for which permission was obtained were included in the corpus.With the exception of a bilingual English-Spanish thesis, all thesis were written in English. The theses are available in PDF format with complex layouts typical to an urbanism project in which text and various types of graphics are combined. @tbl-data gives an overview of the data included in the analysis.

@@

-   corpus D

-   N = `r length(emu_words$word)` words

@@

```{r topic-model-lda-1}
#| include: false

@@

```{r fig-topics-per-doc}

# What is the dominant topic in each document?
N = 20
topic_prop_examples <- theta[1:N,]
colnames(topic_prop_examples) <- topic_names
# TODO change to pivot_longer
vizDataFrame <- melt(cbind(data.frame(topic_prop_examples), 
                           document = factor(rownames(topic_prop_examples))), 
                     variable.name = "topic", id.vars = "document")  
ggplot(data = vizDataFrame, aes(topic, value), ylab = "proportion") + 
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +  
  coord_flip() +
  facet_wrap(~ document, ncol = 5) +
  theme(legend.position = "none")

@@

# Extract topics as mixtures of terms from LDA model
emu_topics <- tidy(emu_lda, matrix = "beta")

# The top n most common words in each topic
emu_top_terms <- get_top_words_per_topic(emu_topics, 20)

@@

```{r}
## Trigrams
emu_trigrams <- emu |> 
  get_ngrams(n = 3, title_col = "title", text_col = "text_raw")

@@

freq_by_rank |> 
  ggplot(aes(rank, `term frequency`, color = title)) +
  geom_line(size = 1.1, alpha = 0.8, show.legend = FALSE) +
  geom_abline(intercept = -1.44, slope = -0.7, color = "gray50", linetype = 2) +
  scale_x_log10() +
  scale_y_log10()
```

```{r}
#| eval: false

## Visualise the top n words with the highest tf-idf values for each of a subset of theses
emu_words_count |> 
  get_tf_idf(title_col = "title", word_col = "word") |> 
  filter(title == unique(title)[1:2]) |>  # Keep only the first 6 theses
  arrange(desc(tf_idf)) |> 
  mutate(word = factor(word, levels = rev(unique(word)))) |> 
  group_by(title) |> 
  top_n(15) |> 
  ungroup() |> 
  ggplot(aes(word, tf_idf, fill = title)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  facet_wrap(~ title,  ncol = 3, scales = "free") +
  coord_flip()
```

@@

tidy_words <- emu |> 
  select(title, text_clean) |> 
  unnest_tokens(word, text_clean) |> 
  add_count(word) |> 
  filter(n >= 50) |> 
  select(-n)

# Summarise word embeddings into document embeddings for modeling purposes
word_matrix <- tidy_words |> 
  count(title, word) |> 
  cast_sparse(title, word, n)
